Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-08-25 article.tex
Section 4 updates: the main structure of the text is in place

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Parallelisation}
\label{sec:parallelisation}
%%--------------------------------------------------------------------------------------------------------------------%%
\hl{- strong scaling: largest mesh, speed vs number of cores: I can use Meluxina}
\hl{- (optional) weak scaling: time vs number of cores as we keep the work-per-core constant}

Comments from \cite{Knoll2004}:
"The first is a scalable implementation, in the sense that time per iteration is reduced in inverse proportion to the number of processors (strong scaling), or that time per iteration is constant as problem size and processor number are scaled proportionally (weak scaling). The second is good per processor performance on contemporary cache-based microprocessors. The third is algorithmic scalability, in the sense that the number of iterations to convergence does not grow with increased numbers of processors (or problem size). The third factor arises because the requirement of a scalable implementation generally forces parameterized changes in the algorithm as the number of processors grows. If the convergence is allowed to degrade, however, the overall execution is not scalable, and this must be countered algorithmically."

>>>>

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Parallelisation}
\label{sec:parallelisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this final analysis, the multi-CPU-core parallel scaling performance of the proposed Jacobian-free Newton Krylov method is examined.
Two types of parallel scaling analyses are performed \cite{Knoll2004}:
\begin{itemize}
	\item Strong scaling study: Strong scaling measures how the execution time of a fixed problem size (fixed number of cells) decreases as the number of CPU cores increases.
	Good strong scaling indicates that an application effectively utilises additional CPU cores without significant overhead.
	In contrast, poor strong scaling suggests that adding more cores does not significantly reduce the execution time, often due to increased communication or synchronisation overhead.

	\item Weak scaling study: Weak scaling measures how the execution time changes as the problem size (the number of cells) and the number of CPU cores increase proportionally, where the problem size per CPU core remains (approximately) constant.
	Good weak scaling indicates that the application can efficiently manage larger workloads with more cores without a significant increase in execution time.
	Poor weak scaling implies that the application struggles to maintain performance as the problem size and core count increase.
\end{itemize}

The results from the strong scaling study are shown in Figure \ref{fig:parallelisation_strong}.
The speedup $S$ is defined as $S = \sfrac{t_1}{t_p}$, where $t_1$ is the clock time on one CPU core and $t_p$ is the clock time on $P$ CPU cores.
In the ideal case, the speedup should double when the number of cores is doubled.
In reality, inter-CPU-core communication reduces the parallel scaling efficiency below the ideal.
\hl{Comment on the results: e.g. fastest, slowest, most/least iterations, best/worst scaling}
\hl{Provide insights, if any: e.g. why is one faster or slower}
\begin{figure}[htbp]
	\centering
	\subfigure[Clock times vs number of CPU cores] {\label{fig:parallel_strong_times} \includegraphics[width=0.2\textwidth]{./figures/parallel_strong_times.pdf}}
	\subfigure[Speedup $S$ vs number of CPU cores] {\label{fig:parallel_strong_speedup} \includegraphics[width=0.2\textwidth]{./figures/parallel_strong_speedup.pdf}}
	\caption{Strong parallel scaling study comparing the performance of three preconditioning strategies: a LU direct solver, and ILU(N) and multigrid iterative solvers}
	\label{fig:parallelisation_strong}
\end{figure}

The weak scaling results are shown in Figure \ref{fig:parallelisation_weak}.
The weak scaling efficiency $\eta_w$ is defined as $\eta_w = \sfrac{t^[w]_1}{t^[w]_p}$, where $t^[w]_1$ is the clock time on one CPU core and $t^[w]_p$ is the clock time on $P$ CPU cores where the problem size per CPU core is constant.
In the ideal case, the efficiency should be unity, but inter-CPU-core communication reduces it.
\hl{Comment on the results: e.g. fastest, slowest, most/least iterations, best/worst scaling}
\hl{Provide insights, if any: e.g. why is one faster or slower}
\begin{figure}[htbp]
	\centering
	\subfigure[Clock times vs number of CPU cores] {\label{fig:parallel_weak_times} \includegraphics[width=0.2\textwidth]{./figures/parallel_weak_times.pdf}}
	\subfigure[Weak scaling efficiency $\eta_w$ vs number of CPU cores] {\label{fig:parallel_weak_speedup} \includegraphics[width=0.2\textwidth]{./figures/parallel_weak_efficiency.pdf}}
	\caption{Weak parallel scaling study comparing the performance of three preconditioning strategies: a LU direct solver, and ILU(N) and multigrid iterative solvers}
	\label{fig:parallelisation_weak}
\end{figure}

%Comments from \cite{Knoll2004}:
%"The first is a scalable implementation, in the sense that time per iteration is reduced in inverse proportion to the number of processors (strong scaling), or that time per iteration is constant as problem size and processor number are scaled proportionally (weak scaling). The second is good per processor performance on contemporary cache-based microprocessors. The third is algorithmic scalability, in the sense that the number of iterations to convergence does not grow with increased numbers of processors (or problem size). The third factor arises because the requirement of a scalable implementation generally forces parameterized changes in the algorithm as the number of processors grows. If the convergence is allowed to degrade, however, the overall execution is not scalable, and this must be countered algorithmically."
